### 2025-05-15 draft.tex
docs: Update draft text

% Created 2025-05-14 Wed 12:29
% Intended LaTeX compiler: pdflatex
\documentclass{article}

@@

\begin{icmlauthorlist}
\icmlauthor{Alan F. Munoz}{broad}
\icmlauthor{Tim Treis}{hh,broad}
\icmlauthor{Alexandr A. Kalinin}{broad}
\icmlauthor{Shatavisha Dasgupta}{broad}
\icmlauthor{Fabian Theis}{hh}
\icmlauthor{Shantanu Singh}{broad}
\icmlauthor{Anne E. Carpenter}{broad}
\end{icmlauthorlist}

@@

\begin{abstract}
Quantifying the contents of objects in images is a common challenge in biological imaging. The most widely used software to acomplish require manual adjustments. Here we introduce our library cp\_measure, which provides programmatic access to the most widespread metrics to convert images and objects into features. We then demonstrate that the features are consistent to the standard ones and showcase tasks for which our tool is more suitable than the alternatives. Our tool opens the door to community-driven  development and expansion of bioimage analysis metrics and pipelines, increasing both the accessibility and reproducibility of the field.
\end{abstract}
\section{Introduction}
\label{sec:org6430daa}
The field of biological imaging leverages data from cell microscopy images to obtain biological insights. Through microscopy techniques we can obtain information from a multitude of sources, such as illumination, fluorescent dyes or even proteins tagged with other proteins that fluoresce, acting as a beacon and providing information on quantity and/or localization.
The main task at the core bioimage analysis is identifying features that convey information of the state of a cell or its contents. This entails combining identified objects with pixel data to quantify their morphology and other information sources, such as intensity distributions.

@@

A CellProfiler build pipeline requires cannot easily integrate with others. Its existing plug-in system is inaccessible for most users and has strong limitations. Building automated pipelines that contain CellProfiler and integrating tools within its components are time and effort-consuming challenges.

@@

As a proof of concept, we showcased that cp\_measure can be used in a workflow to train a classifier for cell populations. We used it to process 433 3D images of astrocytes containing 831 cells \citep{kalinin3DCellNuclear2018}. We then calculated the median value for every feature in a cell and the number of cells, following standard procedures \citep{caicedoDataanalysisStrategiesImagebased2017}. Then we trained a Gradient Boosting classifier to identify which day. With this we identified which features distinguish cells on the later samples and distinguish subpopulations. It is worth noting that there will be some redundancy in the information contained in the cp\_measure features, and thus during cases in which multiple features inform on the same groups of cells Shapley values is likely to only capture one of them.

@@

We plan to add other quality-of-life features. First, add a wrapper to call measurements on the fly, analogous to scikit-image's regionprops function. Secondly, a list of essential measurements for use-cases where speed is more important than using all the features. By lowering the barrier of effort required to integrate cp\_measure into existing pipelines these

>>>>

% Created 2025-05-15 Thu 13:57
% Intended LaTeX compiler: pdflatex
\documentclass{article}

@@

\begin{icmlauthorlist}
\icmlauthor{Alan F. Munoz}{broad}
\icmlauthor{Tim Treis}{hh,broad}
\icmlauthor{Alexandr A. Kalinin}{broad}
\icmlauthor{Shatavisha Dasgupta}{broad}
\icmlauthor{Fabian Theis}{hh}
\icmlauthor{Anne E. Carpenter}{broad}
\icmlauthor{Shantanu Singh}{broad}
\end{icmlauthorlist}

@@

\begin{abstract}
Quantifying the contents of objects in images is a common challenge in biological imaging. The most widely used software to do so require significant manual intervention. Here we introduce our library cp\_measure, which provides programmatic access to the most widespread metrics to convert images and objects into features. We then demonstrate that the features are consistent to the standard ones and showcase tasks for which our tool is more suitable than the alternatives. Our tool opens the door to community-driven  development and expansion of bioimage analysis metrics and pipelines, increasing both the accessibility and reproducibility of the field.
\end{abstract}
\section{Introduction}
\label{sec:org6430daa}
The field of biological imaging leverages data from cell microscopy images to obtain biological insights. Microscopy techniques allows us to acquire information from a multitude of sources, such as illumination, fluorescent dyes or even proteins tagged with other proteins that fluoresce, acting as a beacon and providing information on quantity and/or localization.

The main task at the core bioimage analysis is identifying features that convey information of the state of a cell or its contents. This entails combining identified objects with pixel data to quantify their morphology and other information sources, such as intensity distributions.

@@

A CellProfiler pipeline is does not integrate well with other tools. Its existing plug-in system is inaccessible for most users and has strong limitations. Building automated pipelines that contain CellProfiler and integrating tools within its components are time and effort-consuming challenges.

@@

As a demonstration of its ease of use, we used cp\_measure for featurization in a cell classification workflow. We used it to process 433 3D images of astrocytes containing 831 cells \citep{kalinin3DCellNuclear2018}. We then calculated the median value for every feature in a cell and the number of cells, following standard procedures \citep{caicedoDataanalysisStrategiesImagebased2017}. Then we trained a Gradient Boosting classifier to identify which day. With this we identified which features distinguish cells on the later samples and distinguish subpopulations. It is worth noting that there will be some redundancy in the information contained in the cp\_measure features, and thus during cases in which multiple features inform on similar data subsets.

@@

There is further space for improvement. First, provide a wrapper for all features that masp to scikit-image's regionprops as close as possible. Secondly, a list of essential measurements for use-cases where speed is more important than using all the features. By lowering the barrier of effort required to integrate cp\_measure into existing pipelines these
